---
output: github_document
---

```{r, include = FALSE}
lgr::get_logger("mlr3")$set_threshold("error")
lgr::get_logger("bbotk")$set_threshold("error")
set.seed(1)
options(datatable.print.class = FALSE, datatable.print.keys = FALSE, width = 200)
```

# mlr3hyperband

Package website: [release](https://mlr3hyperband.mlr-org.com/) | [dev](https://mlr3hyperband.mlr-org.com/dev/)

<!-- badges: start -->
[![tic](https://github.com/mlr-org/mlr3hyperband/workflows/tic/badge.svg?branch=main)](https://github.com/mlr-org/mlr3hyperband/actions)
[![CRAN Status](https://www.r-pkg.org/badges/version-ago/mlr3hyperband)](https://cran.r-project.org/package=mlr3hyperband)
[![StackOverflow](https://img.shields.io/badge/stackoverflow-mlr3-orange.svg)](https://stackoverflow.com/questions/tagged/mlr3)
[![Mattermost](https://img.shields.io/badge/chat-mattermost-orange.svg)](https://lmmisld-lmu-stats-slds.srv.mwn.de/mlr_invite/)
<!-- badges: end -->

`mlr3hyperband` extends the [mlr3tuning](https://mlr3tuning.mlr-org.com/) package with various multifidelity optimization methods based on successive halving algorithm.
It currently provides the following optimizers for [bbotk](https://bbotk.mlr-org.com/) and tuner for [mlr3tuning](https://mlr3tuning.mlr-org.com/):

* Successive Halving (`OptimizerSuccessiveHalving` & `TunerSuccessiveHalving`)
* Asynchronous Successive Halving (`OptimizerAsha` & `TunerAsha`)
* Hyperband (`OptimizerSuccessiveHalving` & `TunerSuccessiveHalving`)
* Asynchronous Hyperband (`OptimizerAhb` & `TunerAhb`)

## Resources

* mlr3book chapter on [hyperband](https://mlr3book.mlr-org.com/optimization.html#hyperband)
and [hyperparameter tuning](https://mlr3book.mlr-org.com/optimization.html#tuning).
* The original publications introducing the [successive halving](https://arxiv.org/abs/1502.07943),
[hyperband](https://arxiv.org/abs/1603.06560) and [asha](https://arxiv.org/abs/1810.05934) algorithm.
* Ask questions on [Stackoverflow (tag #mlr3)](https://stackoverflow.com/questions/tagged/mlr3)

## Installation

Install the last release from CRAN:

```{r, eval = FALSE}
install.packages("mlr3hyperband")
```

Install the development version from GitHub:

```{r, eval = FALSE}
remotes::install_github("mlr-org/mlr3hyperband")
```

## Examples

```{r, include = FALSE}
# mute load messages
library(mlr3verse)
library(mlr3hyperband)
library(mlr3learners)
```

```{r}
library(mlr3hyperband)
library(mlr3learners)

# load learner and define search space
learner = lrn("classif.xgboost",
  nrounds = to_tune(p_int(lower = 1, upper = 16, tags = "budget")),
  eta = to_tune(0, 1),
  booster = to_tune()
)

# set parallel backend
future::plan("multisession")

# hyperparameter tuning on the pima indians diabetes data set
instance = tune(
  method = "asha",
  task = tsk("pima"),
  learner = learner,
  resampling = rsmp("cv", folds = 3),
  measures = msr("classif.ce"),
  term_evals = 100
)
```
