---
output: github_document
---

```{r, include = FALSE}
lgr::get_logger("mlr3")$set_threshold("warn")
lgr::get_logger("bbotk")$set_threshold("warn")
set.seed(1)
options(datatable.print.class = FALSE, datatable.print.keys = FALSE, width = 200)
```

# mlr3hyperband

Package website: [release](https://mlr3hyperband.mlr-org.com/) | [dev](https://mlr3hyperband.mlr-org.com/dev/)

<!-- badges: start -->
[![tic](https://github.com/mlr-org/mlr3hyperband/workflows/tic/badge.svg?branch=main)](https://github.com/mlr-org/mlr3hyperband/actions)
[![CRAN Status](https://www.r-pkg.org/badges/version-ago/mlr3hyperband)](https://cran.r-project.org/package=mlr3hyperband)
[![StackOverflow](https://img.shields.io/badge/stackoverflow-mlr3-orange.svg)](https://stackoverflow.com/questions/tagged/mlr3)
[![Mattermost](https://img.shields.io/badge/chat-mattermost-orange.svg)](https://lmmisld-lmu-stats-slds.srv.mwn.de/mlr_invite/)
<!-- badges: end -->

This package provides hyperband tuning for [`mlr3`](https://mlr3.mlr-org.com).

## Installation

Install the last release from CRAN:

```{r, eval = FALSE}
install.packages("mlr3hyperband")
```

Install the development version from GitHub:

```{r, eval = FALSE}
remotes::install_github("mlr-org/mlr3hyperband")
```

## Resources

* mlr3book chapter on [hyperband](https://mlr3book.mlr-org.com/optimization.html#hyperband)
and [hyperparameter tuning](https://mlr3book.mlr-org.com/optimization.html#tuning).
* The original [paper](https://arxiv.org/abs/1603.06560) introducing the hyperband algorithm.

## Hyperband

Hyperband is a budget oriented-procedure, weeding out suboptimally performing configurations early on during their training process aiming at increasing the efficiency of the tuning procedure.
For this, several brackets are constructed with an associated set of configurations for each bracket.
These configuration are initialized by stochastic, often uniform, sampling.
Each bracket is divided into multiple stages, and configurations are evaluated for a increasing budget in each stage.
Note that currently all configurations are trained completely from the beginning, so no online updates to the models are performed.

Different brackets are initialized with different number of configurations, and different budget sizes.
To identify the budget for evaluating hyperband, the user has to specify explicitly which hyperparameter of the learner influences the budget by tagging a single hyperparameter in the parameter set with `"budget"`.
An alternative approach using subsampling and pipelines is described further below.

## Examples

```{r, include = FALSE}
# mute load messages
library(bbotk)
library(mlr3verse)
library(mlr3hyperband)
library(mlr3learners)
```

### Basic

If you are already familiar with `mlr3tuning`, then the only change compared to other tuners is to give a numeric hyperparameter a `budget` tag.
Afterwards, you can handle hyperband like all other tuners.
Originally, hyperband was created with a “natural” learning parameter as the budget parameter in mind, like `nrounds` of the XGBoost learner.

```{r}
library(mlr3verse)
library(mlr3hyperband)
library(mlr3learners)

# define hyperparameter and budget parameter
search_space = ps(
  nrounds = p_int(lower = 1, upper = 16, tags = "budget"),
  eta = p_dbl(lower = 0, upper = 1),
  booster = p_fct(levels = c("gbtree", "gblinear", "dart"))
)

# hyperparameter tuning on the pima indians diabetes data set
instance = tune(
  method = "hyperband",
  task = tsk("pima"),
  learner = lrn("classif.xgboost", eval_metric = "logloss"),
  resampling = rsmp("cv", folds = 3),
  measures = msr("classif.ce"),
  search_space = search_space
)

# best performing hyperparameter configuration
instance$result
```

### Subsampling

Additionally, it is also possible to use `mlr3hyperband` to tune learners that do not have a natural fidelity parameter.
In such a case `mlr3pipelines` can be used to define data subsampling as a preprocessing step.
Then, the `frac` parameter of subsampling, defining the fraction of the training data to be used, can act as the budget parameter.

```{r}
learner = po("subsample") %>>% lrn("classif.rpart")

# define subsampling parameter as budget
search_space = ps(
  classif.rpart.cp = p_dbl(lower = 0.001, upper = 0.1),
  classif.rpart.minsplit = p_int(lower = 1, upper = 10),
  subsample.frac = p_dbl(lower = 0.1, upper = 1, tags = "budget")
)

# hyperparameter tuning on the pima indians diabetes data set
instance = tune(
  method = "hyperband",
  task = tsk("pima"),
  learner = learner,
  resampling = rsmp("cv", folds = 3),
  measures = msr("classif.ce"),
  search_space = search_space
)

# best performing hyperparameter configuration
instance$result
```

### Successive Halving

```{r, message = FALSE}
library(mlr3hyperband)
library(mlr3learners)

# define hyperparameter and budget parameter
search_space = ps(
  nrounds = p_int(lower = 1, upper = 16, tags = "budget"),
  eta = p_dbl(lower = 0, upper = 1),
  booster = p_fct(levels = c("gbtree", "gblinear", "dart"))
)

# hyperparameter tuning on the pima indians diabetes data set
instance = tune(
  method = "successive_halving",
  task = tsk("pima"),
  learner = lrn("classif.xgboost", eval_metric = "logloss"),
  resampling = rsmp("cv", folds = 3),
  measures = msr("classif.ce"),
  search_space = search_space
)

# best performing hyperparameter configuration
instance$result
```

### Quick general-purpose optimization

```{r, message = FALSE}
library(bbotk)
library(mlr3hyperband)

# define hyperparameter and budget parameter
search_space = domain = ps(
  x1 = p_dbl(-5, 10),
  x2 = p_dbl(0, 15),
  fidelity = p_dbl(1e-2, 1, tags = "budget")
)

# modified branin function
objective = ObjectiveRFun$new(
  fun = branin,
  domain = domain,
  codomain = ps(y = p_dbl(tags = "minimize"))
)

# optimize branin function with hyperband
result = bb_optimize(objective, method = "hyperband", search_space = search_space, term_evals = NULL)

# optimized parameters
result$par

# optimal outcome
result$value
```
